Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-07 main.tex
reordering and clean up

This manuscript aims to fill this gap. How to construct potentials for the gradient over triangulated domains is well-documented in the literature. 
We apply the construction principle to the curl and divergence operators, as well as the exterior derivative. 
However, to proceed in the general exterior derivative case, we restrict our efforts to so-called shellable triangulations. 
Only contractible domains can ever admit a shellable triangulation, but having computable upper bounds for such domains is an important stepping stone towards more general situations. 
The class of shellable of shellable triangulations includes practically relevant triangulations. 
For example, local triangulations around a simplex within a larger triangulation (the so-called local patches or stars) are shellable in dimensions two and three. 
Additionally, the study of regularized Poincar\'e and Bogovski\u{i} operators leads to new Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex over bounded convex domains.

@@

% Guerini Savo  
The literature on Poincar\'e--Friedrichs constants and potentials of curls and divergences is less extensive than for potentials of gradients, even over convex domains.
Guerini and Savo~\cite{guerini2004eigenvalue} provide explicit lower bounds for the spectrum of the Hodge--Laplace operator on bounded convex domains with smooth boundary in the Hilbert space case $p=2$.
These bounds are only in terms of the dimension and diameter of the convex domain. 
They thus give upper estimates for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence in particular for the curl and divergence operators.
Additionally, a comparison result shows that the Poincar\'e--Friedrichs constant for the gradient already estimates the corresponding constants for the curl and divergence operators. 
Using a duality argument, this also gives upper estimates of the Poincar\'e--Friedrichs constants for the gradient, curl, and divergence \cye{operators} subject to Dirichlet, tangential, and normal boundary conditions, respectively, along the entire boundary. 
However, these upper bounds are not necessarily optimal. 
Moreover, no results as those of~\cite{guerini2004eigenvalue} are known over bounded convex Lipschitz domains and for general Lebesgue exponents $1 \leq p \leq \infty$.

@@

When the domain is not convex but star-shaped with respect to a ball,
then several estimates for Poincar\'e--Friedrichs constants are known. 
Poincar\'e--Friedrichs inequalities actually hold over merely star-shaped open bounded sets~\cite[Theorem~3.1]{hurri1988poincare}. 
Polynomial interpolation estimates already imply the Poincar\'e--Friedrichs inequality~\cite{brenner2008mathematical,ern2021finite}. 
We pay particular attention to the regularized Poincar\'e and Bogovski\u{i} potential operators for the exterior derivative,
such as those of Costabel and McIntosh~\cite{costabel2010bogovskiui}.
In principle, the operator norms of those potential operators as mapping between Lebesgue spaces of differential forms 
are upper estimates for the Poincar\'e--Friedrichs constants of the domain. 
Here, estimates for the higher-order seminorms of these potentials are available in~\cite{guzman2021estimation}, 
but estimates in Lebesgue norms have not been made explicit in the literature yet, to the best of our knowledge.
We particularly emphasize that these estimates for domains star-shaped with respect to a ball have in common that they rely on upper bounds for the \emph{eccentricity} of the domain.

@@

\subsection{General results}

@@

\subsection{Review: Potentials subject to partial boundary conditions}

@@

\cye{Finally, our main result for convex domains in the exterior calculus setting (Section~\ref{section:intro_EC}) is the construction} of regularized potential operators for the exterior derivative with explicitly bounded operator norms. \cye{This is summarized in Theorem~\ref{theorem:bogovpoinc}.} \mwl{MV: The discussion from before Theorem~\ref{theorem:bogovpoinc} resumed here.} \cye{The arising upper bounds for} the Poincar\'e--Friedrichs constants are proportional to the domain diameter and are independent of the Lebesgue exponent $1 \leq p \leq \infty$, though the \cye{space} dimension and the form degree enter\cye{\sout{s}} the estimate\cye{s}.

\subsection{Combinatorial constants over face-connected triangulated domains}

\subsection{Combinatorial constants over shellable triangulated domains}

\subsection{Main results}

@@

Domains that admit finite triangulations allow for different techniques to attain Poincar\'e--Friedrichs inequalities. 
However, the majority of results only address such inequalities for scalar-valued functions, not for vector fields. 
We review some of the main outcomes. 
Veeser and Verf\"urth provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for finite element stars~\cite{veeser2012poincare}. Naturally, their estimate deteriorates when the shape regularity of the mesh deteriorates.

While we know Poincar\'e--Friedrichs constants over local stars for scalar-valued Sobolev spaces, 
we are not aware of computable estimates for the case of $\bfW^{p}(\curl,\Omega)$ and $\bfW^{p}(\divergence,\Omega)$ over finite element stars. 
In the Hilbert space setting, with $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space with Dirichlet boundary conditions implies, 
via duality, the Poincar\'e--Friedrichs inequality for the divergence operator. However, that easy observation seems restricted to the Hilbert space case $p=2$. 
We also remark that recent work~\cite{gallistl2023computational} addresses eigenvalues for the curl--curl operator over general triangulations,
which gives rise to an estimate for the Poincar\'e--Friedrichs constant when $p=2$. However, the use case of these estimates are sequences of triangulations after sufficiently many levels of refinement, whereas our main application are estimates that work with fixed but small triangulations, such as local stars. 
We are not aware of any estimates for the curl operator, even in the Hilbert space situation.

Towards our main objective of constructing linear potential operators for the gradient, curl, and divergence, satisfying explicitly computable upper bounds, we closely follow the philosophy of the aforementioned sequential approach. Transitioning to vector fields incurs new challenges due to the infinite-dimensional kernels. We overcome these challenges on a special class of triangulated domains, namely domains with so-called \emph{shellable} triangulations. 
\\

\subsection{Objective: sequential approach} 
We want to generalize the sequential construction of the potential operator for the gradient to vector potentials for the curl and divergence. The basic inductive strategy remains the same. We start by constructing, say, a curl potential, over a single simplex. Having already defined a potential operator over a subtriangulation, we then construct a potential over a neighboring simplex. However, the two potentials might not have the same tangential trace along the interface between the subtriangulation and the new simplex. 
% 
To ensure the correct continuity properties, we require stronger conditions on the triangulations than in the case of gradient potentials. We postulate that the intermediate subtriangulations are manifolds with boundary, and that the aforementioned interface is, more specifically, a boundary submanifold of the intermediate subtriangulation and the new simplex. Under those conditions, which are met by so-called \textit{shellable} triangulations, we define a well-posed auxiliary problem for the curl operator with partial tangential boundary conditions. Repeating this procedure exhausts the original triangulation and eventually provides the desired potential.
\\

@@

@@

    The bounds for the Bogovski\u{i} potential operators, Theorem~\ref{theorem:bogovpoinc} imply 
    the existence of $\tilde w \in W^{p}_{0}\Lambda^{k}(\widetilde \Makoto)$
    with $\cartan \tilde w = \cartan \tilde u$ and 
    \begin{align*}
        \| \tilde w \|_{L^{p}(\widetilde \Makoto)}
        &\leq 
        n C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_2(0)) }{ \vol(\widetilde \Makoto) } 
        2
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        \\&\leq 
        2^{n+1} n C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_1(0)) }{ (l+1)/n! } 
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        .
    \end{align*}
    We let $w \in W^{p}\Lambda^{k}(T)$ be defined by $w := \varphi^{-\ast} ( \tilde w_{|\Delta^n} )$.
    By construction, $\cartan w = \cartan u$,
    and $w$ has vanishing trace along $F_{0} \cup \dots \cup F_{l}$.

@@

We first assess the constant in Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative},
which relies on estimates for the Bogovski\u{i} operator over an auxiliary reference domain.
Our experiments over triangles and tetrahedra indicate that we overestimate the constant by several magnitudes.
%
We conjecture that over convex domains $\Omega$, the Poincar\'e--Friedrichs constant of the gradient $\grad : W^{1,p}(\Omega) \rightarrow \bfL^{p}(\Omega)$ is an upper estimate for other Poincar\'e--Friedrichs constants in de~Rham complex. Numerical examples indicate that this holds for $p=2$.

@@

Firstly, we use local Poincar\'e--Friedrichs inequalities over single simplices, subject to various boundary conditions, but these local estimates over simplices (and, more generally, convex domains) likely allow room for improvement. We conjecture that the Guerini--Savo~\cite{guerini2004eigenvalue} estimates, hitherto only known when $p=2$ and over bounded convex domains with smooth boundary, could be extended to the entire range of Lebesgue exponents $1 \leq p \leq \infty$ and convex Lipschitz domains. 
We believe such estimates would yield considerably better bounds on the Poincar\'e--Friedrichs constants than the regularized Poincar\'e operators in the spirit of Costabel and McIntosh~\cite{costabel2010bogovskiui} that we currently employ. Similarly, Poincar\'e--Friedrichs constants subject to boundary conditions allow for upper bounds tighter than what we currently achieve via regularized Bogovski\u{i} operators.

>>>>

This manuscript aims to fill this gap. How to construct potentials for the gradient over triangulated domains is well-documented in the literature. 
We apply the construction principle to the curl and divergence operators, as well as the exterior derivative. 
However, to proceed in the general exterior derivative case, we restrict our efforts to so-called shellable triangulations. 
Only contractible domains can ever admit a shellable triangulation, but having computable upper bounds for such domains is an important stepping stone towards more general situations. 
The class of shellable of shellable triangulations includes practically relevant triangulations. 
For example, local triangulations around a simplex within a larger triangulation (the so-called local patches or stars) are shellable in dimensions two and three. 
Additionally, the study of regularized Poincar\'e and Bogovski\u{\i} operators leads to new Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex over bounded convex domains.

@@

% Guerini Savo  
The literature on Poincar\'e--Friedrichs constants and potentials of curls and divergences is less extensive than for potentials of gradients, even over convex domains.
Guerini and Savo~\cite{guerini2004eigenvalue} provide explicit lower bounds for the spectrum of the Hodge--Laplace operator on bounded convex domains with smooth boundary in the Hilbert space case $p=2$.
These bounds are only in terms of the dimension and diameter of the convex domain. 
They thus give upper estimates for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence in particular for the curl and divergence operators.
Additionally, a comparison result shows that the Poincar\'e--Friedrichs constant for the gradient already estimates the corresponding constants for the curl and divergence operators. 
Using a duality argument, this also gives upper estimates of the Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators subject to Dirichlet, tangential, and normal boundary conditions, respectively, along the entire boundary. 
However, these upper bounds are not necessarily optimal. 
Moreover, no results as those of~\cite{guerini2004eigenvalue} are known over bounded convex Lipschitz domains and for general Lebesgue exponents $1 \leq p \leq \infty$.

@@

When the domain is not convex but star-shaped with respect to a ball,
then several estimates for Poincar\'e--Friedrichs constants are known. 
Poincar\'e--Friedrichs inequalities actually hold over merely star-shaped open bounded sets~\cite[Theorem~3.1]{hurri1988poincare}. 
Polynomial interpolation estimates already imply the Poincar\'e--Friedrichs inequality~\cite{brenner2008mathematical,ern2021finite}. 
We pay particular attention to the regularized Poincar\'e and Bogovski\u{\i} potential operators for the exterior derivative,
such as those of Costabel and McIntosh~\cite{costabel2010bogovskiui}.
In principle, the operator norms of those potential operators as mapping between Lebesgue spaces of differential forms 
are upper estimates for the Poincar\'e--Friedrichs constants of the domain. 
Here, estimates for the higher-order seminorms of these potentials are available in~\cite{guzman2021estimation}, 
but estimates in Lebesgue norms have not been made explicit in the literature yet, to the best of our knowledge.
We particularly emphasize that these estimates for domains star-shaped with respect to a ball have in common that they rely on upper bounds for the \emph{eccentricity} of the domain.

@@

Domains that admit finite triangulations allow for different techniques to attain Poincar\'e--Friedrichs inequalities. 
However, the majority of results only address such inequalities for scalar-valued functions, not for vector fields. 
We review some of the main outcomes.

Veeser and Verf\"urth provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for finite element stars~\cite{veeser2012poincare}. Naturally, their estimate deteriorates when the shape regularity of the mesh deteriorates.

We also remark that recent work~\cite{gallistl2023computational} addresses eigenvalues for the curl--curl operator over general triangulations,
which gives rise to an estimate for the Poincar\'e--Friedrichs constant when $p=2$. However, the use case of these estimates are sequences of triangulations after sufficiently many levels of refinement, whereas our main application are estimates that work with fixed but small triangulations, such as local stars. 
We are not aware of any estimates for the curl operator, even in the Hilbert space situation.

\subsection{Review: Combinatorial constants over face-connected triangulated domains}

\subsection{Review: General results}

@@

\subsection{Objectives: Potentials subject to partial boundary conditions}

@@

\cye{Finally, our main result for convex domains in the exterior calculus setting (Section~\ref{section:intro_EC}) is the construction} of regularized potential operators for the exterior derivative with explicitly bounded operator norms. \cye{This is summarized in Theorem~\ref{theorem:bogovpoinc}.} \mwl{MV: The discussion from before Theorem~\ref{theorem:bogovpoinc} resumed here.} \cye{The arising upper bounds for} the Poincar\'e--Friedrichs constants are proportional to the domain diameter and are independent of the Lebesgue exponent $1 \leq p \leq \infty$, though the \cye{space} dimension and the form degree enter\cye{\sout{s}} the estimate\cye{s}. 
\mwl{Of course, the eccentricity enters the estimate.}

\subsection{Objective: Main results}

@@

\subsection{Objective: General}

While we know Poincar\'e--Friedrichs constants over local stars for scalar-valued Sobolev spaces, 
we are not aware of computable estimates for the case of $\bfW^{p}(\curl,\Omega)$ and $\bfW^{p}(\divergence,\Omega)$ over finite element stars.

In the Hilbert space setting, with $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space with Dirichlet boundary conditions implies, 
via duality, the Poincar\'e--Friedrichs inequality for the divergence operator. However, that easy observation seems restricted to the Hilbert space case $p=2$.

\subsection{Objective: sequential approach} 
Towards our main objective of constructing linear potential operators for the gradient, curl, and divergence, satisfying explicitly computable upper bounds, we closely follow the philosophy of the aforementioned sequential approach. Transitioning to vector fields incurs new challenges due to the infinite-dimensional kernels. We overcome these challenges on a special class of triangulated domains, namely domains with so-called \emph{shellable} triangulations. 
\\

We want to generalize the sequential construction of the potential operator for the gradient to vector potentials for the curl and divergence. The basic inductive strategy remains the same. We start by constructing, say, a curl potential, over a single simplex. Having already defined a potential operator over a subtriangulation, we then construct a potential over a neighboring simplex. However, the two potentials might not have the same tangential trace along the interface between the subtriangulation and the new simplex. 
% 
To ensure the correct continuity properties, we require stronger conditions on the triangulations than in the case of gradient potentials. We postulate that the intermediate subtriangulations are manifolds with boundary, and that the aforementioned interface is, more specifically, a boundary submanifold of the intermediate subtriangulation and the new simplex. Under those conditions, which are met by so-called \textit{shellable} triangulations, we define a well-posed auxiliary problem for the curl operator with partial tangential boundary conditions. Repeating this procedure exhausts the original triangulation and eventually provides the desired potential.
\\

@@

\subsection{Objective: Combinatorial constants over shellable triangulated domains}

@@

    The bounds for the Bogovski\u{\i} potential operators, Theorem~\ref{theorem:bogovpoinc} imply 
    the existence of $\tilde w \in W^{p}_{0}\Lambda^{k}(\widetilde \Makoto)$
    with $\cartan \tilde w = \cartan \tilde u$ and 
    \begin{align*}
        \| \tilde w \|_{L^{p}(\widetilde \Makoto)}
        &\leq 
        n C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_2(0)) }{ \vol(\widetilde \Makoto) } 
        2
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        \\&\leq 
        2^{n+1} n C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_1(0)) }{ (l+1)/n! } 
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        .
    \end{align*}
    We let $w \in W^{p}\Lambda^{k}(T)$ be defined by $w := \varphi^{-\ast} ( \tilde w_{|\Delta^n} )$.
    By construction, $\cartan w = \cartan u$,
    and $w$ has vanishing trace along $F_{0} \cup \dots \cup F_{l}$.

@@

We first assess the constant in Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative},
which relies on estimates for the Bogovski\u{\i} operator over an auxiliary reference domain.
Our experiments over triangles and tetrahedra indicate that we overestimate the constant by several magnitudes.
%
We conjecture that over convex domains $\Omega$, the Poincar\'e--Friedrichs constant of the gradient $\grad : W^{1,p}(\Omega) \rightarrow \bfL^{p}(\Omega)$ is an upper estimate for other Poincar\'e--Friedrichs constants in de~Rham complex. Numerical examples indicate that this holds for $p=2$.

@@

Firstly, we use local Poincar\'e--Friedrichs inequalities over single simplices, subject to various boundary conditions, but these local estimates over simplices (and, more generally, convex domains) likely allow room for improvement. We conjecture that the Guerini--Savo~\cite{guerini2004eigenvalue} estimates, hitherto only known when $p=2$ and over bounded convex domains with smooth boundary, could be extended to the entire range of Lebesgue exponents $1 \leq p \leq \infty$ and convex Lipschitz domains. 
We believe such estimates would yield considerably better bounds on the Poincar\'e--Friedrichs constants than the regularized Poincar\'e operators in the spirit of Costabel and McIntosh~\cite{costabel2010bogovskiui} that we currently employ. Similarly, Poincar\'e--Friedrichs constants subject to boundary conditions allow for upper bounds tighter than what we currently achieve via regularized Bogovski\u{\i} operators.
